```{r message=FALSE}
# library loading
library(dplyr)
library(tidyr)
library(ggplot2)
library(here)
library(scales) # for color gradient
```

```{r}
# file loading for the vignette
old_adf <- readRDS(here("data","budyko_old_whc.rds"))
new_adf <- readRDS(here("data","budyko_new_whc.rds"))
geo <- readRDS(here("data","site_coordinates.rds"))
```

## comparison with normalized AET

```{r}
delta_budyko <- data.frame(sitename = new_adf$sitename)
# just for check
#identical(new_adf$sitename, old_adf$sitename)

delta_budyko$delta_aet <- (old_adf$aet/old_adf$prec_cond)-(new_adf$aet/new_adf$prec_cond)
delta_budyko <- left_join(delta_budyko, geo, by = "sitename")
delta_budyko$delta_whc <- old_adf$whc-new_adf$whc

world <- map_data("world")

europe <- world[world$long < 37 & world$long > -10 & world$lat < 73 & world$lat > 35,]

delta_europe <- delta_budyko[delta_budyko$lon < 37 & delta_budyko$lon > -10 &
                               delta_budyko$lat < 73 & delta_budyko$lat > 35,]

america <- world[world$long < -50 & world$long > -170 & world$lat < 71 & world$lat > 20,]

delta_america <- delta_budyko[delta_budyko$lon < -50 & delta_budyko$lon > -170 &
                               delta_budyko$lat < 71 & delta_budyko$lat > 20,]

asia <- world[world$long < 175 & world$long > 75 & world$lat < 50 & world$lat > -50,]

delta_asia <- delta_budyko[delta_budyko$lon < 175 & delta_budyko$lon > 75 &
                               delta_budyko$lat < 50 & delta_budyko$lat > -50,]

```

## WHC comparison

```{r message=FALSE, warning=FALSE}
# WHC new vs. old 
data.frame(old_WHC = old_adf$whc, 
           new_WHC = new_adf$whc) |>
  ggplot(aes(old_WHC, new_WHC)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  xlim(0, NA) +
  ylim(0, NA)
```



## AET comparison

```{r warning=FALSE}
# AET new vs. old 
data.frame(old_AET = old_adf$aet/old_adf$prec_cond, 
           new_AET = new_adf$aet/new_adf$prec_cond) |>
  ggplot(aes(old_AET, new_AET)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  xlim(0, NA) +
  ylim(0, NA)

```

## Delta WHC over delta AET

```{r message=FALSE, warning=FALSE}

ggplot(delta_budyko, aes(y= delta_aet, x = delta_whc)) + 
  geom_point()
  
```


## Visualize difference in AET and WHC


```{r message=FALSE, warning=FALSE}
ggplot(delta_budyko) +

geom_map(data = world, map = world,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_aet)) +
  scale_color_viridis_c() +
  ggtitle("Delta AET")

```

## delta WHC globally

```{r message=FALSE, warning=FALSE}
ggplot(delta_budyko) +
  geom_map(data = world, map = world,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_whc)) +
  scale_color_viridis_c() +
  ggtitle("Delta WHC")

```

## Delta AET in Europe

```{r message=FALSE, warning=FALSE}
ggplot(delta_europe) +
 geom_map(data = europe, map = europe,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_aet)) +
  scale_color_viridis_c() +
  ggtitle("Delta AET")
```

## Delta WHC in Europe

```{r message=FALSE, warning=FALSE}
ggplot(delta_europe) +
  geom_map(data = europe, map = europe,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_whc)) +
  scale_color_viridis_c() +
  ggtitle("Delta WHC")

```

## Delta WHC and AET combined

```{r message=FALSE, warning=FALSE}
world <- map_data("world")

ggplot(delta_budyko) +

geom_map(data = world, map = world,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_aet, size = delta_whc)) +
  scale_color_viridis_c() +
  ggtitle("Delta AET and WHC")

```

## Deltas in Europe

```{r message=FALSE, warning=FALSE}
ggplot(delta_europe) +
 geom_map(data = europe, map = europe,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_aet,size = delta_whc )) +
  scale_color_viridis_c() +
  ggtitle("Delta AET and WHC")
```

## Deltas in America

```{r message=FALSE, warning=FALSE}
ggplot(delta_america) +
 geom_map(data = america, map = america,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_aet,size = delta_whc )) +
  scale_color_viridis_c() +
  ggtitle("Delta AET and WHC")
```

## Asia

```{r message=FALSE, warning=FALSE}
ggplot(delta_asia) +
 geom_map(data = asia, map = asia,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_aet,size = delta_whc )) +
  scale_color_viridis_c() +
  ggtitle("Delta AET and WHC")
```


## comparison with unnormalised AET

```{r}
delta_budyko <- data.frame(sitename = new_adf$sitename)
# just for check
#identical(new_adf$sitename, old_adf$sitename)

delta_budyko$delta_aet <- old_adf$aet-new_adf$aet
delta_budyko <- left_join(delta_budyko, geo, by = "sitename")
delta_budyko$delta_whc <- old_adf$whc-new_adf$whc


delta_europe <- delta_budyko[delta_budyko$lon < 37 & delta_budyko$lon > -10 &
                               delta_budyko$lat < 73 & delta_budyko$lat > 35,]

delta_america <- delta_budyko[delta_budyko$lon < -50 & delta_budyko$lon > -170 &
                               delta_budyko$lat < 71 & delta_budyko$lat > 20,]

delta_asia <- delta_budyko[delta_budyko$lon < 175 & delta_budyko$lon > 75 &
                               delta_budyko$lat < 50 & delta_budyko$lat > -50,]
```


## AET comparison

```{r warning=FALSE}
# AET new vs. old 
data.frame(old_AET = old_adf$aet, 
           new_AET = new_adf$aet) |>
  ggplot(aes(old_AET, new_AET)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  xlim(0, NA) +
  ylim(0, NA)

```

## Delta WHC over delta AET

```{r message=FALSE, warning=FALSE}

ggplot(delta_budyko, aes(y= delta_aet, x = delta_whc)) + 
  geom_point()
  
```


## Visualize difference in AET and WHC

I'll visualize the difference in AET and WHC globally and then zoomed in Europe due to the high number of sites


```{r message=FALSE, warning=FALSE}
world <- map_data("world")

ggplot(delta_budyko) +

geom_map(data = world, map = world,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_aet)) +
  scale_color_viridis_c() +
  ggtitle("Delta AET")

```


```{r message=FALSE, warning=FALSE}
ggplot(delta_budyko) +
  geom_map(data = world, map = world,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_whc)) +
  scale_color_viridis_c() +
  ggtitle("Delta WHC")

```



```{r message=FALSE, warning=FALSE}
ggplot(delta_europe) +
 geom_map(data = europe, map = europe,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_aet)) +
  scale_color_viridis_c() +
  ggtitle("Delta AET")
```

```{r message=FALSE, warning=FALSE}
ggplot(delta_europe) +
  geom_map(data = europe, map = europe,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_whc)) +
  scale_color_viridis_c() +
  ggtitle("Delta WHC")

```

```{r message=FALSE, warning=FALSE}
world <- map_data("world")

ggplot(delta_budyko) +

geom_map(data = world, map = world,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_aet, size = delta_whc)) +
  scale_color_viridis_c() +
  ggtitle("Delta AET and WHC")

```

## Europe

```{r message=FALSE, warning=FALSE}
ggplot(delta_europe) +
 geom_map(data = europe, map = europe,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_aet,size = delta_whc )) +
  scale_color_viridis_c() +
  ggtitle("Delta AET and WHC")
```

## America

```{r message=FALSE, warning=FALSE}
ggplot(delta_america) +
 geom_map(data = america, map = america,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_aet,size = delta_whc )) +
  scale_color_viridis_c() +
  ggtitle("Delta AET")
```

## Asia

```{r message=FALSE, warning=FALSE}
ggplot(delta_asia) +
 geom_map(data = asia, map = asia,
         aes(long, lat, map_id = region),
         color = "black", fill = "white", size = 0.1) +
  geom_point( aes(lon, lat, color= delta_aet,size = delta_whc )) +
  scale_color_viridis_c() +
  ggtitle("Delta AET and WHC")
```
